Applications of Many-Core Technologies to On-line Event Reconstruction in High Energy Physics Experiments

### 2013-11-13 papers/2013/NSS/NSS.tex
minor adds for the setup and conclusions

ADDITIONAL PARAGRAPH ON THE FERMILAB SETUP.

@@

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an implementation  
using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (8-core), and running the \textsc{opencl} algorithm
on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}

@@

\section{Conclusions}
We have implemented a full version of the CDF SVT tracking algorithm
on GPUs and Intel MIC. We examined a staged approach to using
accelerator cards in a hadron collider trigger application. We have
demonstrated that in this application, significant gains can be
achieved with the `embarrassingly parallel' approach on an Intel MIC
architecture, with the smallest amount of required changes to an
existing serial code base. However, better performance is achieved
with GPUs and a more complete event-level parallelization using
\textsc{cuda}.  We have updated latency studies and shown that for larger packet sizes 
(greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 
since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
small size packets.
%We have updated latency studies and shown that for
%larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
%outperforms P2P, contrary to naive expectation.  
Even at large packet
size, the data transfer takes an appreciable fraction of the total
algorithm time (about 20-25\%).

>>>>

We have an additional setup for testing the \textsc{opencl} implementation of the track 
fitting algorithm. On this setup, we use a 3.07 GHz Intel Core i7 CPU 950, which has four cores 
and up to eight computation threads. We can run the algorithm serially on this CPU (one core), or 
run the \textsc{opencl} algorithm and make use of the multi-core architecture. Additionally, we have an 
AMD Radeon HD 7970 GPU in a PCIe slot in this setup, on which we can also run the \textsc{opencl} algorithm.

@@

\subsection{Data processing in \textsc{opencl}}
We also measure the track fitting algorithm latency in an implementation  
using \textsc{opencl}. Due to limitations on the available space for storing data on the GPU, we only test the 
algorithm on up to 1000 events (\textit{i.e.} about 2 million combinations to fit). In Fig.~\ref{fig:opencl_timing} 
we show the results of algorithm latency measurements for three different modes: running the serial algorithm  
on the CPU (single-core), running the \textsc{opencl} algorithm on the CPU (multi-core), and running the \textsc{opencl} algorithm
on the AMD Radeon GPU. The \textsc{opencl} implementation of the algorithm on the CPU provides a significant speedup---about 
a factor of 5---over running the algorithm serially on the CPU. Running the same \textsc{opencl} algorithm on the GPU provides an 
even greater speedup due to the increased number of cores and parallel threads that can be run, though there is additional latency to copy data 
into and out of the GPU that makes running on the GPU take longer for small numbers of combinations to fit.
\begin{figure}[!t]
  \centering
  \includegraphics[width=0.9\linewidth]{figures/TotalTimeOpencl_1000}
  \caption{The timing of the \textsc{opencl} algorithm as a function of the
    number of track fits. We compare timing running the algorithm serially on a CPU, 
    using \textsc{opencl} on the CPU, and on an AMD GPU.}
  \label{fig:opencl_timing}
\end{figure}

@@

\section{Conclusions}
We have implemented a full version of the CDF SVT tracking algorithm
on GPUs and Intel MIC. We examined a staged approach to using
accelerator cards in a hadron collider trigger application. We have
demonstrated that in this application, significant gains can be
achieved with the `embarrassingly parallel' approach on an Intel MIC
architecture, with the smallest amount of required changes to an
existing serial code base. However, better performance is achieved
with GPUs and a more complete event-level parallelization using
\textsc{cuda}.  Additionally, we have implemented a parallelized algorithm using 
\textsc{opencl} and we have run that algorithm on a multi-core CPU and GPU, showing a boost
in performance over serial CPU computation for even small numbers of events. 
We have also updated latency studies and shown that for larger packet sizes 
(greater than 600 kB), \textsc{cuda}-aware MPI outperforms P2P. This result fully agrees with expectations 
since the P2P mechanism is optimized (and effective) to reduce the transfer latency for 
small size packets.
%We have updated latency studies and shown that for
%larger packet sizes (greater than 600~kB), \textsc{cuda}-aware MPI
%outperforms P2P, contrary to naive expectation.  
Even at large packet
size, the data transfer takes an appreciable fraction of the total
algorithm time (about 20-25\%).
